Child Poverty Interventions

### 2024-05-29 minos/outcomes/child_poverty_reduction_vis.Rmd
Lots of additions and changes to child poverty visualisation scripts

```{r, include=FALSE}
out.path <- here::here('output', 'default_config/')
base.dat <- read_singular_local_out(out.path, 'baseline', 
                                    drop.dead = TRUE,
                                    drop.zero.weight = TRUE)
int.dat <- read_singular_local_out(out.path, 'ChildPovertyReductionRELATIVE', 
                                   drop.dead = TRUE,
                                   drop.zero.weight = TRUE)
int2.dat <- read_singular_local_out(out.path, 'ChildPovertyReductionABSOLUTE', 
                                   drop.dead = TRUE,
                                   drop.zero.weight = TRUE)

#base.dat <- read_agg_subset_batch_out(out.path, 'baseline', drop.dead = TRUE)
#int1.dat <- read_agg_subset_batch_out(out.path, 'ChildPovertyReduction', drop.dead = TRUE)
```

## Threshold

```{r}
base.2021 <- base.dat %>%
  filter(time == 2021)

mean.hh_income <- weighted.mean(base.2021$hh_income, w = base.2021$weight)
rel.pov.threshold <- mean.hh_income * 0.6

print(paste0("Relative poverty threshold in 2020: ", rel.pov.threshold))
rm(base.2021, mean.hh_income, rel.pov.threshold)
```

```{r}
threshold <- base.dat %>%
  select(time, weight, hh_income) %>%
  group_by(time) %>%
  summarise(poverty_threshold = (weighted.mean(hh_income, w = weight) * 0.6))

ggplot(data = threshold, aes(x = time, y = poverty_threshold)) +
  geom_col() +
  geom_hline(yintercept = threshold$poverty_threshold[threshold$time == 2020], linetype='dashed') +
  labs(title = 'Relative poverty threshold over time') +
  xlab('Year') +
  ylab('Threshold')
```

## Intervention

In this intervention, we will randomly select households to uplift in each year until we uplift the number of children required for that year.

### Boosted Sample

hh_boosted <- int.dat %>%
    select(hidp, time, income_boosted, boost_amount, nkids) %>%
    filter(income_boosted == 'True') %>%
    group_by(time) %>%
    summarise(n_boosted = length(unique(hidp)),
              total_boost = sum(boost_amount),
              kids_boosted = sum(nkids))

sample_information <- function(int.dat, target.year) {
  hh_boosted <- int.dat %>%
    select(hidp, time, income_boosted, boost_amount, nkids) %>%
    filter(income_boosted == 'True') %>%
    group_by(time) %>%
    summarise(n_boosted = length(unique(hidp)),
              total_boost = sum(boost_amount),
              kids_boosted = sum(nkids))

@@

  print(p1)
  print(p2)
  print(p3)
  print(p4)
}

sample_information(int.dat, target.year = 2030)
```

### SF_12_MCS

#### Full Population

First lets look at the effect of the intervention on the entire sample.

```{r}
full.sample.change.SF12 <- function(base.dat, int.dat) {
  base.all.SF12 <- base.dat %>%
    select(time, weight, SF_12) %>%
    group_by(time) %>%
    summarise(base.SF12 = weighted.mean(SF_12, w = weight))

@@

  ggplot(combined, aes(x = time, y = diff)) +
    geom_line() +
    geom_hline(yintercept = 0, linetype='dashed') +
    labs(title = 'Relative change in SF12') +
    xlab('Year') +
    ylab('Difference')
}

full.sample.change.SF12(base.dat, int.dat)
```

#### Families with children

Now focused only on families with children.

```{r}
all.families.change.SF12 <- function(base.dat, int.dat) {
  base.all.SF12 <- base.dat %>%
    select(time, weight, SF_12, nkids) %>%
    filter(nkids > 0) %>%
    group_by(time) %>%
    summarise(base.SF12 = weighted.mean(SF_12, w = weight))

@@

  ggplot(combined, aes(x = time, y = diff)) +
    geom_line() +
    geom_hline(yintercept = 0, linetype='dashed') +
    labs(title = 'Relative change in SF12') +
    xlab('Year') +
    ylab('Difference')
}
all.families.change.SF12(base.dat, int.dat)
```

#### Families w/ children under poverty line

Now only households with children under the poverty line.

```{r}
poverty.families.change.SF12 <- function(base.dat, int.dat, threshold) {
  base.thresh <- merge(base.dat, threshold, by = 'time')
  int.thresh <- merge(int.dat, threshold, by = 'time')

@@

  ggplot(combined, aes(x = time, y = diff)) +
    geom_line() +
    geom_hline(yintercept = 0, linetype='dashed') +
    labs(title = 'Relative change in SF12') +
    xlab('Year') +
    ylab('Difference')
}

poverty.families.change.SF12(base.dat, int.dat, threshold)
```

This plot is perhaps a bit disingenuous, as the closer we get to 2030 the fewer households are actually under the poverty line. We might achieve what we want better by looking only at the boosted population.

#### Treatment on the Treated

```{r}
treatment.on.treated.change.SF12 <- function(base.dat, int.dat) {
  boosted.byyear <- int.dat %>%
    filter(income_boosted == 'True')

  base.all.SF12 <- base.dat %>%
    select(time, weight, SF_12, nkids, pidp) %>%
    filter(pidp %in% boosted.byyear$pidp) %>%
    group_by(time) %>%
    summarise(base.SF12 = weighted.mean(SF_12, w = weight))

  int.all.SF12 <- int.dat %>%
    select(time, weight, SF_12, nkids, pidp) %>%
    filter(pidp %in% boosted.byyear$pidp) %>%
    group_by(time) %>%
    summarise(int.SF12 = weighted.mean(SF_12, w = weight))

@@

  ggplot(combined, aes(x = time, y = diff)) +
    geom_line() +
    geom_hline(yintercept = 0, linetype='dashed') +
    labs(title = 'Relative change in SF12') +
    xlab('Year') +
    ylab('Difference')
}

treatment.on.treated.change.SF12(base.dat, int.dat)
```

## Absolute Poverty

### Boosted Sample

```{r}
sample_information(int2.dat, target.year = 2030)
```

### SF_12_MCS

#### Full Population

```{r}
full.sample.change.SF12(base.dat, int2.dat)
```

#### Families with children

```{r}
all.families.change.SF12(base.dat, int.dat)
```

#### Families w/ children under poverty line

```{r}
poverty.families.change.SF12(base.dat, int.dat, threshold)
```

#### Treatment on the Treated

```{r}
treatment.on.treated.change.SF12(base.dat, int.dat)
```

@@

```{r}
base.spag <- base.dat %>%
  select(pidp, age, time, hh_income) %>%
  mutate(scen = 'baseline')
int.spag <- int.dat %>%
  select(pidp, age, time, hh_income) %>%
  mutate(scen = 'relative')
int2.spag <- int2.dat %>%
  select(pidp, age, time, hh_income) %>%
  mutate(scen = 'absolute')

@@

ggplot(base.dat, aes(x = hh_income, y = time, group = time, color = 'blue', alpha = 0.3)) +
  geom_density_ridges() +
  geom_density_ridges(data = int.dat, aes(x = hh_income, y = time, group = time, color = 'red')) +
  theme(legend.position = 'none') +
  labs(title = 'Relative Poverty Intervention')

ggplot(base.dat, aes(x = hh_income, y = time, group = time, color = 'blue', alpha = 0.3)) +
  geom_density_ridges() +
  geom_density_ridges(data = int2.dat, aes(x = hh_income, y = time, group = time, color = 'red')) +
  theme(legend.position = 'none') +
  labs(title = 'Absolute Poverty Intervention')

@@

>>>>

```{r, include=FALSE}
out.path <- here::here('output', 'default_config/')
base.dat <- read_singular_local_out(out.path, 'baseline', 
                                    drop.dead = TRUE,
                                    drop.zero.weight = TRUE)
# int1.rel <- read_singular_local_out(out.path, 'ChildPovertyReductionRELATIVE',
#                                    drop.dead = TRUE,
#                                    drop.zero.weight = TRUE)
int2.rel_fixed <- read_singular_local_out(out.path, 'ChildPovertyReductionRELATIVE_2',
                                   drop.dead = TRUE,
                                   drop.zero.weight = TRUE)
int3.abs <- read_singular_local_out(out.path, 'ChildPovertyReductionABSOLUTE',
                                   drop.dead = TRUE,
                                   drop.zero.weight = TRUE)

#base.dat <- read_agg_subset_batch_out(out.path, 'baseline', drop.dead = TRUE)
#int1.dat <- read_agg_subset_batch_out(out.path, 'ChildPovertyReduction', drop.dead = TRUE)

int2.rel_fixed <- int2.rel_fixed #%>%
  #filter(age < 91)

int3.abs <- int3.abs #%>%
  #filter(age < 91)
```

# Functions

```{r sample_info}
sample_information <- function(int.dat, target.year) {
  hh_boosted <- int.dat %>%
    select(hidp, time, income_boosted, boost_amount, nkids) %>%
    filter(income_boosted == 'True') %>%
    group_by(time) %>%
    summarise(n_boosted = length(unique(hidp)),
              total_boost = sum(boost_amount),
              kids_boosted = sum(nkids))

@@

  print(p1)
  print(p2)
  print(p3)
  print(p4)
}
```

```{r sf_12_change}
full.sample.change.SF12 <- function(base.dat, int.dat) {
  base.all.SF12 <- base.dat %>%
    select(time, weight, SF_12) %>%
    group_by(time) %>%
    summarise(base.SF12 = weighted.mean(SF_12, w = weight))

@@

  ggplot(combined, aes(x = time, y = diff)) +
    geom_line() +
    geom_hline(yintercept = 0, linetype='dashed') +
    labs(title = 'Relative change in SF12', subtitle = 'Full Population') +
    xlab('Year') +
    ylab('Difference')
}

all.families.change.SF12 <- function(base.dat, int.dat) {
  base.all.SF12 <- base.dat %>%
    select(time, weight, SF_12, nkids) %>%
    filter(nkids > 0) %>%
    group_by(time) %>%
    summarise(base.SF12 = weighted.mean(SF_12, w = weight))

@@

  ggplot(combined, aes(x = time, y = diff)) +
    geom_line() +
    geom_hline(yintercept = 0, linetype='dashed') +
    labs(title = 'Relative change in SF12', subtitle = 'Families') +
    xlab('Year') +
    ylab('Difference')
}

poverty.families.change.SF12.variable <- function(base.dat, int.dat, threshold) {
  base.thresh <- merge(base.dat, threshold, by = 'time')
  int.thresh <- merge(int.dat, threshold, by = 'time')

@@

  ggplot(combined, aes(x = time, y = diff)) +
    geom_line() +
    geom_hline(yintercept = 0, linetype='dashed') +
    labs(title = 'Relative change in SF12', subtitle = 'Families in Poverty') +
    xlab('Year') +
    ylab('Difference')
}

poverty.families.change.SF12 <- function(base.dat, int.dat, threshold_val) {

  base.all.SF12 <- base.dat %>%
    select(time, weight, SF_12, nkids, hh_income) %>%
    filter(nkids > 0) %>%
    filter(hh_income < threshold_val) %>%
    group_by(time) %>%
    summarise(base.SF12 = weighted.mean(SF_12, w = weight))

  int.all.SF12 <- int.dat %>%
    select(time, weight, SF_12, nkids, hh_income) %>%
    filter(nkids > 0) %>%
    filter(hh_income < threshold_val) %>%
    group_by(time) %>%
    summarise(int.SF12 = weighted.mean(SF_12, w = weight))

@@

  ggplot(combined, aes(x = time, y = diff)) +
    geom_line() +
    geom_hline(yintercept = 0, linetype='dashed') +
    labs(title = 'Relative change in SF12', subtitle = 'Families in Poverty') +
    xlab('Year') +
    ylab('Difference')
}

treatment.on.treated.change.SF12 <- function(base.dat, int.dat) {
  boosted.pidps <- int.dat %>%
    filter(income_boosted == 'True') %>%
    select(pidp, time)

  base.all.SF12 <- base.dat %>%
    select(time, weight, SF_12, nkids, pidp) %>%
    inner_join(boosted.pidps, by = c("pidp", "time")) %>%
    group_by(time) %>%
    summarise(base.SF12 = weighted.mean(SF_12, w = weight))

  int.all.SF12 <- int.dat %>%
    select(time, weight, SF_12, nkids, pidp) %>%
    inner_join(boosted.pidps, by = c("pidp", "time")) %>%
    group_by(time) %>%
    summarise(int.SF12 = weighted.mean(SF_12, w = weight))

  ggplot(combined, aes(x = time, y = diff)) +
    geom_line() +
    geom_hline(yintercept = 0, linetype='dashed') +
    labs(title = 'Relative change in SF12', subtitle = 'Treatment on the Treated') +
    xlab('Year') +
    ylab('Difference')
}
```

# Absolute Poverty

## Boosted Sample

```{r}
sample_information(int3.abs, target.year = 2030)
```

## SF_12_MCS

```{r}
full.sample.change.SF12(base.dat, int3.abs)

all.families.change.SF12(base.dat, int3.abs)

poverty.families.change.SF12(base.dat, int3.abs, threshold_val = 955.54)

treatment.on.treated.change.SF12(base.dat, int3.abs)
```

# Relative Poverty (FIXED THRESHOLD)

## Boosted Sample

```{r}
sample_information(int2.rel_fixed, target.year = 2030)
```

## SF_12_MCS

```{r}
full.sample.change.SF12(base.dat, int2.rel_fixed)

all.families.change.SF12(base.dat, int2.rel_fixed)

poverty.families.change.SF12(base.dat, int2.rel_fixed, threshold_val = 1247.71)

treatment.on.treated.change.SF12(base.dat, int2.rel_fixed)
```

@@

```{r}
base.spag <- base.dat %>%
  select(pidp, age, time, hh_income) %>%
  mutate(scen = 'baseline')
int.spag <- int2.rel_fixed %>%
  select(pidp, age, time, hh_income) %>%
  mutate(scen = 'relative')
int2.spag <- int3.abs %>%
  select(pidp, age, time, hh_income) %>%
  mutate(scen = 'absolute')

@@

ggplot(base.dat, aes(x = hh_income, y = time, group = time, color = 'blue', alpha = 0.3)) +
  geom_density_ridges() +
  geom_density_ridges(data = int2.rel_fixed, aes(x = hh_income, y = time, group = time, color = 'red')) +
  theme(legend.position = 'none') +
  labs(title = 'Relative Poverty Intervention (FIXED)')

ggplot(base.dat, aes(x = hh_income, y = time, group = time, color = 'blue', alpha = 0.3)) +
  geom_density_ridges() +
  geom_density_ridges(data = int3.abs, aes(x = hh_income, y = time, group = time, color = 'red')) +
  theme(legend.position = 'none') +
  labs(title = 'Absolute Poverty Intervention')

@@

# Relative Poverty Variable Threshold (DEFUNKT)

```{r}
# base.2021 <- base.dat %>%
#   filter(time == 2021)
# 
# mean.hh_income <- weighted.mean(base.2021$hh_income, w = base.2021$weight)
# rel.pov.threshold <- mean.hh_income * 0.6
# 
# print(paste0("Relative poverty threshold in 2020: ", rel.pov.threshold))
# rm(base.2021, mean.hh_income, rel.pov.threshold)
```

```{r}
# threshold <- base.dat %>%
#   select(time, weight, hh_income) %>%
#   group_by(time) %>%
#   summarise(poverty_threshold = (weighted.mean(hh_income, w = weight) * 0.6))
# 
# ggplot(data = threshold, aes(x = time, y = poverty_threshold)) +
#   geom_col() +
#   geom_hline(yintercept = threshold$poverty_threshold[threshold$time == 2020], linetype='dashed') +
#   labs(title = 'Relative poverty threshold over time') +
#   xlab('Year') +
#   ylab('Threshold')
```

## Boosted Sample

```{r}
# sample_information(int1.rel, target.year = 2030)
```

## SF_12_MCS

```{r}
# full.sample.change.SF12(base.dat, int1.rel)
# all.families.change.SF12(base.dat, int1.rel)
# poverty.families.change.SF12.variable(base.dat, int1.rel, threshold)
# treatment.on.treated.change.SF12(base.dat, int1.rel)
```
